---
title: "MTH 2520 - R Programming"
subtitle: "Using Python in R Markdown"
author: "TYPE YOUR NAME HERE"
output:
  word_document: 
      reference_docx: "../../WordStyleRef-01.docx"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,comment="")
library(rmarkdown, tinytex)

```

***

**Some Examples of Python code** 

You need to have Python installed on your computer

Install and load the package: reticulate 

Set the path where your Python executable is located: 

```{r}
library(reticulate)
use_python("C:/")
```

Click on 'Insert' button above and choose Python to insert Python code chunk. 

Here are some basic Python chunks: 

Compute square roots:

```{python}
# Use Python to calculate the square root
# Note: change this value for a different result
num = 8 

# To take the input from the user:
#num = float(input('Enter a number: '))

num_sqrt = num ** 0.5

print('The square root of %0.3f is %0.3f'%(num ,num_sqrt))
```

Compute roots of the quadratic equation:

```{python}
# Solve the quadratic equation ax**2 + bx + c = 0
# import complex math module
import cmath
a = 1
b = 4
c = 5

# calculate the discriminant
d = b**2 - 4*a*c

# find two solutions
sol1 = (-b-cmath.sqrt(d))/(2*a)
sol2 = (-b+cmath.sqrt(d))/(2*a)

print('The solution are {0} and {1}'.format(sol1,sol2))

```

Draw a Sierpinski triangle:

<!--This chunk uses turtle graphics and draws the graph in a separate window. I haven't figured out how to draw the picture in the document. I played with some commands to save the picture as .eps, but it took an eternity. Maybe some of you Python experts can fix this for me. -->

```{python}
import turtle
PROGNAME = 'Sierpinski Triangle'
#Credits: This code was written by editing the code from http://www.lpb-riannetrujillo.com/blog/python-fractal/

myPen = turtle.Turtle()
myPen.ht()
myPen.speed(5)
myPen.pencolor('blue')

points = [[-175,-125],[0,175],[175,-125]] #size of triangle

def getMid(p1,p2):
    return ( (p1[0]+p2[0]) / 2, (p1[1] + p2[1]) / 2) #find midpoint

def triangle(points,depth):

    myPen.up()
    myPen.goto(points[0][0],points[0][1])
    myPen.down()
    myPen.goto(points[1][0],points[1][1])
    myPen.goto(points[2][0],points[2][1])
    myPen.goto(points[0][0],points[0][1])

    if depth>0:
        triangle([points[0],
                        getMid(points[0], points[1]),
                        getMid(points[0], points[2])],
                   depth-1)
        triangle([points[1],
                        getMid(points[0], points[1]),
                        getMid(points[1], points[2])],
                   depth-1)
        triangle([points[2],
                         getMid(points[2], points[1]),
                         getMid(points[0], points[2])],
                   depth-1)


triangle(points,4)
```
